### 2022-10-02 manuscript.tex
Merge pull request #274 from SORTEE-Github-Hackathon/RCO-edits-october [ci skip]
This build is based on
https://github.com/SORTEE-Github-Hackathon/manuscript/commit/67184a6d4709f90ee2da7a7f13a18ae8b930330f.

This commit was created by the following CI build and job:
https://github.com/SORTEE-Github-Hackathon/manuscript/commit/67184a6d4709f90ee2da7a7f13a18ae8b930330f/checks
https://github.com/SORTEE-Github-Hackathon/manuscript/actions/runs/3169645123

This manuscript
(\href{https://SORTEE-Github-Hackathon.github.io/manuscript/v/ec11faeb0bef6e3bab5ff066c8b7786218ca8fe5/}{permalink})
was automatically generated
from \href{https://github.com/SORTEE-Github-Hackathon/manuscript/tree/ec11faeb0bef6e3bab5ff066c8b7786218ca8fe5}{SORTEE-Github-Hackathon/manuscript@ec11fae}
on September 22, 2022.

@@

The expansive GitHub user-community and numerous GitHub resources have boosted its popularity\textsuperscript{\protect\hyperlink{ref-RVetqmsg}{7},\protect\hyperlink{ref-kEX5dgzK}{9}--\protect\hyperlink{ref-u5aEVE4B}{11}}.
Nevertheless, although multiple articles have encouraged researchers in EEB to adopt GitHub as part of their research process\textsuperscript{\protect\hyperlink{ref-10ghgV3S8}{3},\protect\hyperlink{ref-3DKwn1sY}{12}}, its use is still not widespread.
First-time users without formal training in information technology may face steep learning curves because GitHub and its features have been centered on collaboration for code development in information systems\textsuperscript{\protect\hyperlink{ref-139b0pSGc}{13}}.
Moreover, domain-specific resources providing tractable examples and practical guidance for researchers in EEB on GitHub are scarce (but see \url{https://ourcodingclub.github.io}; \url{https://www.openscapes.org}).
Widespread adoption of GitHub for collaborating on research tasks can ultimately enable EEB researchers to spend less time on creating novel processes for collaboration and more time on their research\textsuperscript{\protect\hyperlink{ref-ydrk01SR}{14}}.
More importantly, expanding the availability of data and code management standards -- of which GitHub is one increasingly important component -- makes research more reproducible and collaborative\textsuperscript{\protect\hyperlink{ref-13QX8XU3J}{15},\protect\hyperlink{ref-pq2Tv1BC}{16}}.

@@

An EEB research compendium includes all computational materials related to research production, including data, code for analyses and protocols.
Having copies of these files safely stored is essential to protect against accidental modifications or deletions.
Many researchers begin using GitHub to store (or backup) their research compendium\textsuperscript{\protect\hyperlink{ref-MwwMapRG}{17}} to a centralized, readily-available remote server (see \protect\hyperlink{definitions}{Box 1}).
A centralized research compendium stored in a version-controlled repository has the advantages of facilitating collaboration, integrating data and code archiving services, allowing file versions to be accessed and restored, and contributes to open science (see sections below).

GitHub limits committed file sizes to 100 Mb (megabytes\textsuperscript{\protect\hyperlink{ref-1Co6ZZjF1}{18}}), which can make it challenging for centralizing research compendia containing larger file sizes.
Users may still version large files using Git Large File Storage\textsuperscript{\protect\hyperlink{ref-11GtZ7icJ}{19}} text pointers, but may have to rely on external file storage alternatives (such as local or cloud-hosting).

@@

Laboratory notebooks help researchers track their research notes, methods, policies and protocols\textsuperscript{\protect\hyperlink{ref-wwHxTOtm}{20}}.
Virtual laboratory notebooks can be stored in GitHub repositories and provide the benefits of simultaneous, centralized and selective access, and allows for the easy update of policies and experiment protocols\textsuperscript{\protect\hyperlink{ref-10V7x4H4l}{21}}.
Researchers have been increasingly using GitHub to maintain versions and share digital laboratory notebooks across many fields\textsuperscript{\protect\hyperlink{ref-10ghgV3S8}{3}}, including applied ecology (\emph{e.g.}, \url{https://scheuerell-lab.github.io/lab-book}), biogeography and global change biology (\emph{e.g.}, \url{https://github.com/HuckleyLab/how_we_work}), and microbial ecology (\emph{e.g.}, \url{https://github.com/CarBBAS/uqam-guide}).

@@

Modern research in ecology and evolution is highly collaborative, bringing together multidisciplinary teams from various institutions\textsuperscript{\protect\hyperlink{ref-1HhzKAC1K}{22}}.
On GitHub, collaborators can share feedback, brainstorm ideas, and troubleshoot problems (Figure \ref{fig:github-diagram}).
Project management can happen via three GitHub repository features: ``Issues'', ``Discussion'' and ``Projects''.
Github Issues allow for discrete tasks and sub-tasks to be identified, assigned to team members, and categorized with custom labels.
Github Discussions serve as a message board for conversation.
Finally, GitHub Projects integrate issues and pull requests on automated spreadsheets and project boards, providing users with real-time tracking of project priorities and status\textsuperscript{\protect\hyperlink{ref-RhBKe0MG}{23}}.
Scripts, commit messages, and pull requests can be linked directly to issues, discussions, and projects providing a clear record of project workflow.
For instance, one can create a discussion on a project repository to decide on which method to apply for analyzing biodiversity data, and then an issue can be created with the purpose of establishing the steps and responsibilities covering from data formatting, statistical analyses, until the figure generation and issue resolution {[}\emph{e.g.}, see issues and discussions in the sPlotOpen {[}doi:10.1111/geb.13346{]} manuscript repository (\url{https://github.com/fmsabatini/sPlotOpen_Manuscript/issues?q=is\%3Aissue+is\%3Aclosed}){]}.
Using GitHub for all project-related conversation and planning, rather than email or messaging tools, makes it easier to keep track of progress throughout the lifespan of a project.
Unlike emails and messages which can get lost as more new tasks arise, GitHub issues are intentionally closed by repository administrators hiding the issue from view (closed issues remain accessible but not immediately visible).
Project management in GitHub can also be integrated with third-party applications, such as Zenhub (\url{https://www.zenhub.com}) or Slack (\url{https://slack.github.com}).
ZenHub allows for the enhanced visualization and organization of repositories and their issues, while the GitHub for Slack integration allows notifications from GitHub events to be sent directly to users or group channels\textsuperscript{\protect\hyperlink{ref-1GRWGlDWy}{24}}.

@@

Making presentations, syllabi and other course materials can be done through most major high-level programming languages such as R, with \texttt{RMarkdown}\textsuperscript{\protect\hyperlink{ref-QqMezOMg}{25}}, Python, with \texttt{python-ppt} (\url{https://python-pptx.readthedocs.io}), and Julia, with \texttt{Remark.jl} (\url{https://juliapackages.com/p/remark}), and be version-controlled and stored in GitHub.
Once content is made, hosting a course website can be done through GitHub Pages\textsuperscript{\protect\hyperlink{ref-MDQfMwCW}{26}} (\emph{e.g.}, \url{https://github.com/topics/course-website}).
This way, course content can be available to enrolled students, as well as a global pool of learners and teachers interested in the material.
Content can then be delivered via the course website, and/or a GitHub Organization with, for example, template repositories for assignments.
Student submissions are perhaps the most challenging component, but the new GitHub Classroom tool (\url{https://classroom.github.com}) allows instructors to host private assignments to be submitted collaboratively or individually as code or PDF files, and even build autograding tests.
Although time-consuming to establish, using these features can integrate learning version control and GitHub with the learning course content, and thus boost students' feelings of self-efficacy and confidence\textsuperscript{\protect\hyperlink{ref-dqrFjoSb}{27}}.

@@

Personal or laboratory websites can improve the sharing of research findings, build online presence, and increase coordination of research efforts\textsuperscript{\protect\hyperlink{ref-HiIPSSHV}{28}}.
Despite researchers in ecology and evolution generally lacking experience in building or hosting webpages, many tools have been developed to help this process.
Static websites can now be easily built using independent software and languages, such as Jekyll (\url{https://jekyllrb.com}), Hugo (\url{https://gohugo.io}), Quarto (\url{https://quarto.org}), and wowchemy (\url{https://wowchemy.com}), or with the help of dependencies in the programming languages scientists commonly use, such as the \texttt{distill} (\url{https://github.com/rstudio/distill}) and \texttt{blogdown} (\url{https://bookdown.org/yihui/blogdown}) R packages.
The resulting files from static websites can then be hosted in repositories, from which one can activate GitHub's Pages (\url{https://pages.github.com}) feature, allowing for the direct live hosting of HyperText Markup Language (HTML), Cascading Style Sheets (CSS) and JavaScript files within a free github.io domain.

Website templates are readily available on GitHub (\url{https://github.com/topics/website-template}) or in user-specific repositories, which can be forked (\protect\hyperlink{definitions}{Box 1}) and customized\textsuperscript{\protect\hyperlink{ref-MXxgZJ45}{29}}.
Aside from free hosting services, GitHub Pages also allows websites to be autogenerated and instantly updated following content modifications\textsuperscript{\protect\hyperlink{ref-RVetqmsg}{7}}.
Creating and hosting websites on GitHub pages can be time-consuming, as this process usually requires more knowledge in web design than out-of-the-box platforms (\emph{e.g.}, Wix, Weebly, Google Sites).
However, free hosting, widely available template customization, and versioning are strong advantages over the alternatives.

@@

Government, funding agencies, and publishers exercise rigorous open access data policies and mandates\textsuperscript{\protect\hyperlink{ref-PLmDFZrm}{30},\protect\hyperlink{ref-1Hcf13Q0k}{31}}.
However, code and data sharing may be met by individual reluctance, temporary embargoes, or partially prevented by privacy and confidentiality reasons\textsuperscript{\protect\hyperlink{ref-1Ch6LSHef}{32}--\protect\hyperlink{ref-1CzUZwyU2}{34}}.
Still, data deposition and ensuring its availability can amplify the outreach of published studies\textsuperscript{\protect\hyperlink{ref-666HppfO}{35}}, increase citation rates\textsuperscript{\protect\hyperlink{ref-1CcAUn3Lu}{36}}, and among many other reasons, enable the reproducibility and robustness of scientific advances\textsuperscript{\protect\hyperlink{ref-4LaijDIZ}{37}--\protect\hyperlink{ref-1HZdsK5Kn}{39}}.
While public repositories on GitHub make it easy to store and share data files, they are not considered long-term repositories for research materials.
This is because GitHub, a for-profit company, does not have long-term data availability guarantees, allowing users to delete or make repositories private after publication.
Also, GitHub does not issue Digital Object Identifiers (DOI) for content uploaded to their servers.
DOIs are persistent and citable unique alpha-numeric identifiers assigned to digitally stored research materials.
Because of this, scientists sharing code and data through GitHub are strongly encouraged to independently submit their research materials to long-term data archives (\emph{e.g.}, Zenodo, Figshare, Dryad, OSF\textsuperscript{\protect\hyperlink{ref-10ghgV3S8}{3},\protect\hyperlink{ref-kEX5dgzK}{9},\protect\hyperlink{ref-1Du6fzB8g}{40}}; \protect\hyperlink{tbl:compare}{Table 1}).
Some of these options (Zenodo, Figshare and OSF) integrate with GitHub, allowing project, code, and data releases (\protect\hyperlink{definitions}{Box 1}) to be archived with versioned, citable DOIs.
Linking GitHub repositories with a DOI helps research become findable, properly cited, and can ensure long-term stability\textsuperscript{\protect\hyperlink{ref-iIEKCTLU}{41}}.
This strategy has been increasingly adopted in numerous studies in ecology and evolution (\emph{e.g.},\textsuperscript{\protect\hyperlink{ref-GQj3c17f}{42}};\textsuperscript{\protect\hyperlink{ref-bZNn2hbh}{43}};\textsuperscript{\protect\hyperlink{ref-ZI1OqZNr}{44}}).

An important aspect of making code and data citable and reusable is to add an appropriate license to protect intellectual property.
Code published without a license is under exclusive copyright (by default), protecting it from copy, distribution, and modifications.
One may grant specific rights to their code for reuse by adding licensing files and specifications within GitHub repositories\textsuperscript{\protect\hyperlink{ref-TOsASkn5}{45}}.
The Choose a License (\url{https://choosealicense.com/non-software/}) website offers further guidance on the licenses available for research and creative products.
For example, Creative Commons (CC; \url{https://creativecommons.org/licenses/}) licenses can specify that shared code is intended for a specific analysis.
A CC BY 4.0 license specifies that any code (or other creative products) must be appropriately credited to its original author when distributed, adapted or reused.

@@

Researchers can asynchronously communicate and individually or collaboratively work in GitHub.
By forking, branching, and cloning repositories (\protect\hyperlink{definitions}{Box 1}), researchers can simultaneously work on different issues of the same research project to then later merge changes to the main branch project with pull requests.
Explicit project organization and increased communication in GitHub Issues, Discussions, or within pull requests can help with project development and with potential merge conflicts due to users simultaneously working on the same sections\textsuperscript{\protect\hyperlink{ref-1BJcvyTmV}{46}}.
Moreover, version control\textsuperscript{\protect\hyperlink{ref-K7nbP1Ty}{47}} and commit reversal features allow researchers to track and progress without worrying about irreparably modifying someone else's work.
By enabling more comprehensive remote collaboration, GitHub encourages the exchange of ideas among researchers at different institutions and in different countries, which can serve to improve the quality of the research itself by providing open access to data and code.
In academic settings, GitHub can also facilitate interactions between research advisors and advisees, providing a platform for students or other trainees to share in-progress code, and flag specific challenges or questions for their supervisors or mentors (\protect\hyperlink{tbl:roles}{Table 2}).
Periodic code review\textsuperscript{\protect\hyperlink{ref-hm9PaCLD}{48}} can also help advisors identify errors early in the process, and inform further training and mentorship to fill gaps in skills.

@@

We wrote this manuscript using Manubot\textsuperscript{\protect\hyperlink{ref-Re6Eg2va}{49}}, a modifiable workflow implemented in GitHub to automatically render manuscripts and automate bibliographical tasks\textsuperscript{\protect\hyperlink{ref-YuJbg3zO}{50}}.
Manubot uses GitHub's automation workflow, GitHub Actions, to combine and convert individual Markdown files into a single LaTeX document, which can then be converted to a Word or PDF document, and displayed as a webpage.
Citations and bibliographic references are automatically managed with citable persistent identifiers (\emph{e.g.}, DOIs, PubMed IDs, ISBNs, URLs).
The resulting manuscript can be rendered with document templates and citation style language formatting to meet journal formatting requirements.
Every change made to the manuscript triggers its rendering, so that updates are readily displayed and made publicly available.
Additional GitHub Actions can be integrated with Manubot, such as ones creating figures or generating tables (\emph{e.g.}, \url{https://github.com/SORTEE-Github-Hackathon/manuscript/tree/main/.github/workflows}).

@@

GitHub can also assist reviewers during the peer review process.
If the code associated with a manuscript is made available at the time of submission (\emph{e.g.}, as a link to a GitHub repository within the Data Availability Statement), peer reviewers may be able to offer more comprehensive suggestions on the code and written materials, potentially recognizing errors before publication.
Certain journals or software development communities require submitted work or research code to be hosted on GitHub and their review processes make use of GitHub Issues (\emph{e.g.}, rOpenSci (\url{https://ropensci.org/software-review/}), Journal of Open Source Software (\url{https://joss.readthedocs.io/en/latest/submitting.html})).
rOpenSci's efforts have resulted in many well-used R packages for ecology research including \texttt{rfishbase}\textsuperscript{\protect\hyperlink{ref-Xsdcv6q}{51}} and \texttt{taxize}\textsuperscript{\protect\hyperlink{ref-FVBWKkZu}{52}}.

@@

Scientific publications often omit part of their intellectual and computational workflows, including the treatment of raw data and analytical steps (\emph{e.g.}, model assumption testing).
Publishing data and reproducible workflows along with manuscripts can provide readers with all details about analytical steps and enable reproducing research experiments and results\textsuperscript{\protect\hyperlink{ref-NOgBWVAr}{53}}.
In addition to storing data and code, GitHub repositories can provide a time-stamped (version controlled) preregistration of research plans and hypotheses.

Conventional research practices typically separate tasks among collaborators (\emph{i.e.}, data entry, analysis, writing).
It is common that coauthors discuss, but do not actively verify, edit, or execute research tasks that are not their main responsibility.
GitHub can serve as a tool for open and tractable research development.
Collaborators can directly interact with code and data, inspect for errors, and potentially identify scientific misconduct prior to manuscript submission (\emph{e.g.},\textsuperscript{\protect\hyperlink{ref-18PTmKJkq}{54}}).
Collaborators and readers are better positioned to discover erroneous or questionable findings if they have complete and transparent access to projects.

@@

Projects in ecology and evolution often involve graduate students, research assistants, and post-doctoral fellows, who hold limited-term positions\textsuperscript{\protect\hyperlink{ref-D4C4k4ak}{55}}.
Without clear plans on project continuity, the research code and data management upkeep tends to fall off as researchers move on to new projects or to other institutions.
Additionally, code and data can be difficult to access when kept only on personal devices\textsuperscript{\protect\hyperlink{ref-19kmNxiHc}{56}}.

GitHub can facilitate project continuity among research code by making code handover easier\textsuperscript{\protect\hyperlink{ref-4ny1onB0}{8},\protect\hyperlink{ref-D4C4k4ak}{55}}.
Through version control, the history of code and data from projects in ecology and evolution can be tracked accessible to future laboratory members and collaborators\textsuperscript{\protect\hyperlink{ref-3DKwn1sY}{12}}.
Repositories and organizations can have designated data and code owners (or more appropriate, stewards; see\textsuperscript{\protect\hyperlink{ref-iIEKCTLU}{41}},\textsuperscript{\protect\hyperlink{ref-s91uGRZ2}{57}}), who can also can change through time allowing for the transition of code between research cohorts.
Other project collaborators can contribute to repository design and development, and their active involvement can both aid authors ability to act as guarantors of the project, and the clarity and reproducibility of the project for future users.
In (Figure \ref{fig:github-diagram}), we highlight several elements of good repository structure, and the various ways that contributors may interact with them.

>>>>

This manuscript
(\href{https://SORTEE-Github-Hackathon.github.io/manuscript/v/67184a6d4709f90ee2da7a7f13a18ae8b930330f/}{permalink})
was automatically generated
from \href{https://github.com/SORTEE-Github-Hackathon/manuscript/tree/67184a6d4709f90ee2da7a7f13a18ae8b930330f}{SORTEE-Github-Hackathon/manuscript@67184a6}
on October 2, 2022.

@@

The expansive GitHub user-community and numerous GitHub resources have boosted its popularity\textsuperscript{\protect\hyperlink{ref-RVetqmsg}{7},\protect\hyperlink{ref-kEX5dgzK}{9}--\protect\hyperlink{ref-u5aEVE4B}{11}}.
Nevertheless, although multiple articles have encouraged researchers in EEB to adopt GitHub as part of their research process\textsuperscript{\protect\hyperlink{ref-10ghgV3S8}{3},\protect\hyperlink{ref-3DKwn1sY}{12}}, its use is still not widespread.
First-time users without formal training in information technology may face steep learning curves because GitHub and its features have been centered on collaboration for code development in information systems\textsuperscript{\protect\hyperlink{ref-139b0pSGc}{13}}.
Moreover, domain-specific resources providing tractable examples and practical guidance for researchers in EEB on GitHub are scarce (but see \url{https://ourcodingclub.github.io}; \url{https://www.openscapes.org},\textsuperscript{\protect\hyperlink{ref-lJAgyhYq}{14}}).
Widespread adoption of GitHub for collaborating on research tasks can ultimately enable EEB researchers to spend less time on creating novel processes for collaboration and more time on their research\textsuperscript{\protect\hyperlink{ref-ydrk01SR}{15}}.
More importantly, expanding the availability of data and code management standards -- of which GitHub is one increasingly important component -- makes research more reproducible and collaborative\textsuperscript{\protect\hyperlink{ref-13QX8XU3J}{16},\protect\hyperlink{ref-pq2Tv1BC}{17}}.

@@

An EEB research compendium includes all computational materials related to research production, including data, code for analyses and protocols.
Having copies of these files safely stored is essential to protect against accidental modifications or deletions.
Many researchers begin using GitHub to store (or backup) their research compendium\textsuperscript{\protect\hyperlink{ref-MwwMapRG}{18}} to a centralized, readily-available remote server (see \protect\hyperlink{definitions}{Box 1}).
A centralized research compendium stored in a version-controlled repository has the advantages of facilitating collaboration, integrating data and code archiving services, allowing file versions to be accessed and restored, and contributes to open science (see sections below).

GitHub limits committed file sizes to 100 Mb (megabytes\textsuperscript{\protect\hyperlink{ref-1Co6ZZjF1}{19}}), which can make it challenging for centralizing research compendia containing larger file sizes.
Users may still version large files using Git Large File Storage\textsuperscript{\protect\hyperlink{ref-11GtZ7icJ}{20}} text pointers, but may have to rely on external file storage alternatives (such as local or cloud-hosting).

@@

Laboratory notebooks help researchers track their research notes, methods, policies and protocols\textsuperscript{\protect\hyperlink{ref-wwHxTOtm}{21}}.
Virtual laboratory notebooks can be stored in GitHub repositories and provide the benefits of simultaneous, centralized and selective access, and allows for the easy update of policies and experiment protocols\textsuperscript{\protect\hyperlink{ref-10V7x4H4l}{22}}.
Researchers have been increasingly using GitHub to maintain versions and share digital laboratory notebooks across many fields\textsuperscript{\protect\hyperlink{ref-10ghgV3S8}{3}}, including applied ecology (\emph{e.g.}, \url{https://scheuerell-lab.github.io/lab-book}), biogeography and global change biology (\emph{e.g.}, \url{https://github.com/HuckleyLab/how_we_work}), and microbial ecology (\emph{e.g.}, \url{https://github.com/CarBBAS/uqam-guide}).

@@

Modern research in ecology and evolution is highly collaborative, bringing together multidisciplinary teams from various institutions\textsuperscript{\protect\hyperlink{ref-1HhzKAC1K}{23}}.
On GitHub, collaborators can share feedback, brainstorm ideas, and troubleshoot problems (Figure \ref{fig:github-diagram}).
Project management can happen via three GitHub repository features: ``Issues'', ``Discussion'' and ``Projects''.
Github Issues allow for discrete tasks and sub-tasks to be identified, assigned to team members, and categorized with custom labels.
Github Discussions serve as a message board for conversation.
Finally, GitHub Projects integrate issues and pull requests on automated spreadsheets and project boards, providing users with real-time tracking of project priorities and status\textsuperscript{\protect\hyperlink{ref-RhBKe0MG}{24}}.
Scripts, commit messages, and pull requests can be linked directly to issues, discussions, and projects providing a clear record of project workflow.
For instance, one can create a discussion on a project repository to decide which method to apply for biodiversity data analysis.
Then, an issue can be created to establish steps and responsibilities including data formatting, statistical analyses, figure generation, and issue resolution {[}\emph{e.g.}, see issues and discussions in the sPlotOpen {[}doi:10.1111/geb.13346{]} manuscript repository (\url{https://github.com/fmsabatini/sPlotOpen_Manuscript/issues?q=is\%3Aissue+is\%3Aclosed}){]}.
Using GitHub for all project-related conversation and planning, rather than email or messaging tools, makes it easier to keep track of progress throughout the lifespan of a project.
Unlike emails and messages which can get lost as more new tasks arise, GitHub issues are intentionally closed by repository administrators hiding the issue from view (closed issues remain accessible but not immediately visible).
Project management in GitHub can also be integrated with third-party applications, such as Zenhub (\url{https://www.zenhub.com}) or Slack (\url{https://slack.github.com}).
Zenhub allows for the enhanced visualization and organization of repositories and their issues, while the GitHub for Slack integration allows notifications from GitHub events to be sent directly to users or group channels\textsuperscript{\protect\hyperlink{ref-1GRWGlDWy}{25}}.

@@

Making presentations, syllabi and other course materials can be done through most major high-level programming languages such as R, with \texttt{RMarkdown}\textsuperscript{\protect\hyperlink{ref-QqMezOMg}{26}}, Python, with \texttt{python-ppt} (\url{https://python-pptx.readthedocs.io}), and Julia, with \texttt{Remark.jl} (\url{https://juliapackages.com/p/remark}), and be version-controlled and stored in GitHub.
Once content is made, hosting a course website can be done through GitHub Pages\textsuperscript{\protect\hyperlink{ref-MDQfMwCW}{27}} (\emph{e.g.}, \url{https://github.com/topics/course-website}).
This way, course content can be available to enrolled students, as well as a global pool of learners and teachers interested in the material.
Content can then be delivered via the course website, and/or a GitHub Organization with, for example, template repositories for assignments.
Student submissions are perhaps the most challenging component, but the new GitHub Classroom tool (\url{https://classroom.github.com}) allows instructors to host private assignments to be submitted collaboratively or individually as code or PDF files, and even build autograding tests.
Although time-consuming to establish, using these features can integrate learning version control and GitHub with the learning course content, and thus boost students' feelings of self-efficacy and confidence\textsuperscript{\protect\hyperlink{ref-dqrFjoSb}{28}}.

@@

Personal or laboratory websites can improve the sharing of research findings, build online presence, and increase coordination of research efforts\textsuperscript{\protect\hyperlink{ref-HiIPSSHV}{29}}.
Despite researchers in ecology and evolution generally lacking experience in building or hosting webpages, many tools have been developed to help this process.
Static websites can now be easily built using independent software and languages, such as Jekyll (\url{https://jekyllrb.com}), Hugo (\url{https://gohugo.io}), Quarto (\url{https://quarto.org}), and wowchemy (\url{https://wowchemy.com}), or with the help of dependencies in the programming languages scientists commonly use, such as the \texttt{distill} (\url{https://github.com/rstudio/distill}) and \texttt{blogdown} (\url{https://bookdown.org/yihui/blogdown}) R packages.
The resulting files from static websites can then be hosted in repositories, from which one can activate GitHub's Pages (\url{https://pages.github.com}) feature, allowing for the direct live hosting of HyperText Markup Language (HTML), Cascading Style Sheets (CSS) and JavaScript files within a free github.io domain.

Website templates are readily available on GitHub (\url{https://github.com/topics/website-template}) or in user-specific repositories, which can be forked (\protect\hyperlink{definitions}{Box 1}) and customized\textsuperscript{\protect\hyperlink{ref-MXxgZJ45}{30}}.
Aside from free hosting services, GitHub Pages also allows websites to be autogenerated and instantly updated following content modifications\textsuperscript{\protect\hyperlink{ref-RVetqmsg}{7}}.
Creating and hosting websites on GitHub pages can be time-consuming, as this process usually requires more knowledge in web design than out-of-the-box platforms (\emph{e.g.}, Wix, Weebly, Google Sites).
However, free hosting, widely available template customization, and versioning are strong advantages over the alternatives.

@@

Government, funding agencies, and publishers exercise rigorous open access data policies and mandates\textsuperscript{\protect\hyperlink{ref-PLmDFZrm}{31},\protect\hyperlink{ref-1Hcf13Q0k}{32}}.
However, code and data sharing may be met by individual reluctance, temporary embargoes, or partially prevented by privacy and confidentiality reasons\textsuperscript{\protect\hyperlink{ref-1Ch6LSHef}{33}--\protect\hyperlink{ref-1CzUZwyU2}{35}}.
Still, data deposition and ensuring its availability can amplify the outreach of published studies\textsuperscript{\protect\hyperlink{ref-666HppfO}{36}}, increase citation rates\textsuperscript{\protect\hyperlink{ref-1CcAUn3Lu}{37}}, and among many other reasons, enable the reproducibility and robustness of scientific advances\textsuperscript{\protect\hyperlink{ref-4LaijDIZ}{38}--\protect\hyperlink{ref-1HZdsK5Kn}{40}}.
While public repositories on GitHub make it easy to store and share data files, they are not considered long-term repositories for research materials.
This is because GitHub, a for-profit company, does not have long-term data availability guarantees, allowing users to delete or make repositories private after publication.
Also, GitHub does not issue Digital Object Identifiers (DOI) for content uploaded to their servers.
DOIs are persistent and citable unique alpha-numeric identifiers assigned to digitally stored research materials.
Because of this, scientists sharing code and data through GitHub are strongly encouraged to independently submit their research materials to long-term data archives (\emph{e.g.}, Zenodo, Figshare, Dryad, OSF\textsuperscript{\protect\hyperlink{ref-10ghgV3S8}{3},\protect\hyperlink{ref-kEX5dgzK}{9},\protect\hyperlink{ref-1Du6fzB8g}{41}}; \protect\hyperlink{tbl:compare}{Table 1}).
Some of these options (Zenodo, Figshare and OSF) integrate with GitHub, allowing project, code, and data releases (\protect\hyperlink{definitions}{Box 1}) to be archived with versioned, citable DOIs.
Linking GitHub repositories with a DOI helps research become findable, properly cited, and can ensure long-term stability\textsuperscript{\protect\hyperlink{ref-iIEKCTLU}{42}}.
This strategy has been increasingly adopted in numerous studies in ecology and evolution (\emph{e.g.},\textsuperscript{\protect\hyperlink{ref-GQj3c17f}{43}};\textsuperscript{\protect\hyperlink{ref-bZNn2hbh}{44}};\textsuperscript{\protect\hyperlink{ref-ZI1OqZNr}{45}}).

An important aspect of making code and data citable and reusable is to add an appropriate license to protect intellectual property.
Code published without a license is under exclusive copyright (by default), protecting it from copy, distribution, and modifications.
One may grant specific rights to their code for reuse by adding licensing files and specifications within GitHub repositories\textsuperscript{\protect\hyperlink{ref-TOsASkn5}{46}}.
The Choose a License (\url{https://choosealicense.com/non-software/}) website offers further guidance on the licenses available for research and creative products.
For example, Creative Commons (CC; \url{https://creativecommons.org/licenses/}) licenses can specify that shared code is intended for a specific analysis.
A CC BY 4.0 license specifies that any code (or other creative products) must be appropriately credited to its original author when distributed, adapted or reused.

@@

Researchers can asynchronously communicate and individually or collaboratively work in GitHub.
By forking, branching, and cloning repositories (\protect\hyperlink{definitions}{Box 1}), researchers can simultaneously work on different issues of the same research project to then later merge changes to the main branch project with pull requests.
Explicit project organization and increased communication in GitHub Issues, Discussions, or within pull requests can help with project development and with potential merge conflicts due to users simultaneously working on the same sections\textsuperscript{\protect\hyperlink{ref-1BJcvyTmV}{47}}.
Moreover, version control\textsuperscript{\protect\hyperlink{ref-K7nbP1Ty}{48}} and commit reversal features allow researchers to track and progress without worrying about irreparably modifying someone else's work.
By enabling more comprehensive remote collaboration, GitHub encourages the exchange of ideas among researchers at different institutions and in different countries, which can serve to improve the quality of the research itself by providing open access to data and code.
In academic settings, GitHub can also facilitate interactions between research advisors and advisees, providing a platform for students or other trainees to share in-progress code, and flag specific challenges or questions for their supervisors or mentors (\protect\hyperlink{tbl:roles}{Table 2}).
Periodic code review\textsuperscript{\protect\hyperlink{ref-hm9PaCLD}{49}} can also help advisors identify errors early in the process, and inform further training and mentorship to fill gaps in skills.

@@

We wrote this manuscript using Manubot\textsuperscript{\protect\hyperlink{ref-Re6Eg2va}{50}}, a modifiable workflow implemented in GitHub to automatically render manuscripts and automate bibliographical tasks\textsuperscript{\protect\hyperlink{ref-YuJbg3zO}{51}}.
Manubot uses GitHub's automation workflow, GitHub Actions, to combine and convert individual Markdown files into a single LaTeX document, which can then be converted to a Word or PDF document, and displayed as a webpage.
Citations and bibliographic references are automatically managed with citable persistent identifiers (\emph{e.g.}, DOIs, PubMed IDs, ISBNs, URLs).
The resulting manuscript can be rendered with document templates and citation style language formatting to meet journal formatting requirements.
Every change made to the manuscript triggers its rendering, so that updates are readily displayed and made publicly available.
Additional GitHub Actions can be integrated with Manubot, such as ones creating figures or generating tables (\emph{e.g.}, \url{https://github.com/SORTEE-Github-Hackathon/manuscript/tree/main/.github/workflows}).

@@

GitHub can also assist reviewers during the peer review process.
If the code associated with a manuscript is made available at the time of submission (\emph{e.g.}, as a link to a GitHub repository within the Data Availability Statement), peer reviewers may be able to offer more comprehensive suggestions on the code and written materials, potentially recognizing errors before publication.
Certain journals or software development communities require submitted work or research code to be hosted on GitHub and their review processes make use of GitHub Issues (\emph{e.g.}, rOpenSci (\url{https://ropensci.org/software-review/}), Journal of Open Source Software (\url{https://joss.readthedocs.io/en/latest/submitting.html})).
rOpenSci's efforts have resulted in many well-used R packages for ecology research including \texttt{rfishbase}\textsuperscript{\protect\hyperlink{ref-Xsdcv6q}{52}} and \texttt{taxize}\textsuperscript{\protect\hyperlink{ref-FVBWKkZu}{53}}.

@@

Scientific publications often omit part of their intellectual and computational workflows, including the treatment of raw data and analytical steps (\emph{e.g.}, model assumption testing).
Publishing data and reproducible workflows along with manuscripts can provide readers with all details about analytical steps and enable reproducing research experiments and results\textsuperscript{\protect\hyperlink{ref-NOgBWVAr}{54}}.
In addition to storing data and code, GitHub repositories can provide a time-stamped (version controlled) preregistration of research plans and hypotheses.

Conventional research practices typically separate tasks among collaborators (\emph{i.e.}, data entry, analysis, writing).
It is common that coauthors discuss, but do not actively verify, edit, or execute research tasks that are not their main responsibility.
GitHub can serve as a tool for open and tractable research development.
Collaborators can directly interact with code and data, inspect for errors, and potentially identify scientific misconduct prior to manuscript submission (\emph{e.g.},\textsuperscript{\protect\hyperlink{ref-18PTmKJkq}{55}}).
Collaborators and readers are better positioned to discover erroneous or questionable findings if they have complete and transparent access to projects.

@@

Projects in ecology and evolution often involve graduate students, research assistants, and post-doctoral fellows, who hold limited-term positions\textsuperscript{\protect\hyperlink{ref-D4C4k4ak}{56}}.
Without clear plans on project continuity, the research code and data management upkeep tends to fall off as researchers move on to new projects or to other institutions.
Additionally, code and data can be difficult to access when kept only on personal devices\textsuperscript{\protect\hyperlink{ref-19kmNxiHc}{57}}.

GitHub can facilitate project continuity among research code by making code handover easier\textsuperscript{\protect\hyperlink{ref-4ny1onB0}{8},\protect\hyperlink{ref-D4C4k4ak}{56}}.
Through version control, the history of code and data from projects in ecology and evolution can be tracked accessible to future laboratory members and collaborators\textsuperscript{\protect\hyperlink{ref-3DKwn1sY}{12}}.
Repositories and organizations can have designated data and code owners (or more appropriate, stewards; see\textsuperscript{\protect\hyperlink{ref-iIEKCTLU}{42}},\textsuperscript{\protect\hyperlink{ref-s91uGRZ2}{58}}), who can also can change through time allowing for the transition of code between research cohorts.
Other project collaborators can contribute to repository design and development, and their active involvement can both aid authors ability to act as guarantors of the project, and the clarity and reproducibility of the project for future users.
In (Figure \ref{fig:github-diagram}), we highlight several elements of good repository structure, and the various ways that contributors may interact with them.
